Removal only

### 2020-05-26 analysis/figure-4.Rmd
working on removal, gonads have missing data

---
title: "Fig 4 and Supple figs"
output: md_document
---

@@

knitr::opts_chunk$set(echo = TRUE, fig.path = '../figures/')
```

# Figure 1. Experimental design, tSNE analysis, PCA, and prolactin

@@

# prep for tsne

hyptsne <- subsetmaketsne("hypothalamus", allmaniplevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", allmaniplevels, sexlevels)
gontsne <- subsetmaketsne("gonads", allmaniplevels, sexlevels)

@@

  df <- df %>% mutate(earlylate = fct_collapse(treatment, 
                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
                                         "inc.d9", "lay", "bldg"),
                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
                                         "m.n2", "extend", "n5")),
                extint = fct_collapse(treatment, 
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
                                  chicks = c( "early", "hatch", "extend", "n5"),
                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg")))
  df$extint <- factor(df$extint, levels = levelsextint)
  return(df)
}

hyptsne <- addgroupings(hyptsne)
pittsne <- addgroupings(pittsne)
gontsne <- addgroupings(gontsne)

@@

  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2),
                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
                treatment =  factor(treatment, levels = alllevels),
                earlylate = fct_collapse(treatment, 
                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
                                         "inc.d9", "lay", "bldg"),
                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
                                         "m.n2", "extend", "n5")),
                extint = fct_collapse(treatment, 
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong", "lay"),
                                  chicks = c( "early", "hatch", "extend", "n5"),
                                  loss = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"))) %>%
    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  %>%
    dplyr::filter(tissue == whichtissue) 
  return(candidates)
}

hypvsd <- getcandidatevsdmanip(candidategenes, "hypothalamus")
pitvsd <- getcandidatevsdmanip(candidategenes, "pituitary")
gonvsd <- getcandidatevsdmanip(candidategenes, "gonad")
candidatevsd <- rbind(hypvsd, pitvsd)
candidatevsd <- rbind(candidatevsd, gonvsd)
head(candidatevsd)

@@

plotcandidatetiming <- function(df, whichgene, whichtissue){

  dfslim <- df %>%
    dplyr::filter(gene  == whichgene, treatment %in% timing) 
  print(head(dfslim))

  p <- dfslim %>%
    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    theme_B3() +
    theme(legend.position = "none",
          #axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          plot.subtitle = element_text(face = "italic")) +
    scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
    labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
    facet_wrap(~sex)  +
    geom_signif(comparisons = list(c( "inc.d9", "early"),
                                   c( "inc.d17", "prolong"),
                                   c( "hatch", "prolong"),
                                   c( "hatch", "extend")),
                map_signif_level=TRUE,
                textsize = 3, family = 'Helvetica',
                vjust = 1.5, size = 0)
  return(p)

@@

a <- plotcandidatetiming(hypvsd, "DRD1", "hypothalamus")
b <- plotcandidatetiming(pitvsd, "PRL", "pituitary")
c <- plotcandidatetiming(gonvsd, "CRHBP", "gonads")

>>>>

---
title: "Removal only"
output: md_document
---

@@

knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
                      fig.path = '../figures/', fig.width = 7.25, fig.height = 7.25)
```

# Figure 4. What is the effect of removal

@@

# prep for tsne
hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)

@@

  df <- df %>% mutate(earlylate = fct_collapse(treatment, 
                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
                                         "inc.d9",  "bldg"),
                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
                                         "m.n2", "extend", "n5"),
                                  reference = "control"),
                extint = fct_collapse(treatment, 
                                  nest = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"),
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong"),
                                  chicks = c("early", "hatch", "extend", "n5"),
                                  reference = "control")) 
  return(df)
}

hyptsne <- addgroupings(hyptsne)
pittsne <- addgroupings(pittsne)
gontsne <- addgroupings(gontsne)
head(gontsne)

t1 <- plottsneelipsev2(hyptsne, hyptsne$treatment, allcolors) +  labs(subtitle = "Hypothalamus ") +
  facet_wrap(~sex) + theme(axis.title.x = element_blank())
t2 <- plottsneelipsev2(hyptsne, hyptsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
  facet_wrap(~sex) + theme(axis.title = element_blank())
t3 <- plottsneelipsev2(hyptsne, hyptsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
  facet_wrap(~sex) + theme(axis.title = element_blank())
t123 <- plot_grid(t1, t2, t3, nrow = 1, rel_widths = c(1.1,1,1))

t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(subtitle = "Pituitary ") +
  facet_wrap(~sex) + theme(axis.title.x = element_blank(), strip.text = element_blank())
t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
  facet_wrap(~sex) + theme(axis.title = element_blank(), strip.text = element_blank())
t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
  facet_wrap(~sex) + theme(axis.title = element_blank(), strip.text = element_blank())
t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))

t7 <- plottsneelipsev2(gontsne, gontsne$treatment, allcolors) +  labs(subtitle = "Gonads ") +
  facet_wrap(~sex) + theme(strip.text = element_blank())
t8 <- plottsneelipsev2(gontsne, gontsne$extint, colorhypothesis) +  labs(subtitle = "  ")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
t9 <- plottsneelipsev2(gontsne, gontsne$earlylate, colorhypothesis) +  labs(subtitle = "  ")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank(), strip.text = element_blank())
t789 <- plot_grid(t7, t8, t9, nrow = 1)

tsneplots <- plot_grid(t123, t456,t789, ncol = 1, rel_widths = c(1.1,1,1))
tsneplots
```

@@

  candidates  <- allvsd %>%
    filter(gene %in% whichgenes) %>%
    dplyr::mutate(sextissue = sapply(strsplit(file_name, '_vsd.csv'), "[", 1),
                  sextissue = sapply(strsplit(sextissue, '../results/DEseq2/hypothesis/'), "[", 2),
                  sex = sapply(strsplit(sextissue, '\\_'), "[", 1),
                tissue = sapply(strsplit(sextissue, '\\_'), "[", 2),
                treatment = sapply(strsplit(samples, '\\_'), "[", 4),
                treatment = sapply(strsplit(treatment, '.NYNO'), "[", 1),
                treatment = recode(treatment, "m.hatch" = "m.n2", "extend.hatch" = "m.n2",
                                                   "inc.prolong" = "prolong", "m.inc.d8" = "early"),
                treatment =  factor(treatment, levels = alllevels),
                earlylate = fct_collapse(treatment, 
                                  early = c("early", "m.inc.d3", "m.inc.d9", "inc.d3", 
                                         "inc.d9",  "bldg"),
                                  late = c("inc.d17",   "m.inc.d17", "prolong" ,  "hatch" , 
                                         "m.n2", "extend", "n5"),
                                  reference = "control"),
                extint = fct_collapse(treatment, 
                                  nest = c("m.inc.d3", "m.inc.d9", "m.n2", "m.inc.d17", "bldg"),
                                  eggs = c("inc.d3", "inc.d9", "inc.d17", "prolong"),
                                  chicks = c("early", "hatch", "extend", "n5"),
                                  reference = "control"))  %>%
    dplyr::select(gene, counts, sex, tissue, treatment, earlylate, extint, samples)  
  return(candidates)
}

candidatevsd <- getcandidatevsdmanip(candidategenes)
hypvsd <- candidatevsd %>% filter(tissue %in% "hypothalamus") %>% drop_na()
pitvsd <- candidatevsd %>% filter(tissue %in% "pituitary")%>% drop_na()
gonvsd <- candidatevsd %>% filter(tissue %in% "gonad")%>% drop_na()

@@

plotcandidateremoval <- function(df, whichgene, whichtissue){

  p <- df %>%
    dplyr::filter(gene  == whichgene, treatment %in% removallevels)  %>%
    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    theme_B3() +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          plot.subtitle = element_text(face = "italic")) +
    scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
    labs(title = paste(whichtissue, sep = " "), subtitle = whichgene, y = "gene expression") +
    #facet_wrap(~sex)  +
    geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
                                   c( "inc.d9", "m.inc.d9"),
                                   c( "inc.d17", "m.inc.d17"),
                                   c( "hatch", "m.n2")),
                map_signif_level=TRUE,
                textsize = 3, family = 'Helvetica',
                vjust = 1.5, size = 0.5)
  return(p)

@@

a <- plotcandidateremoval(hypvsd, "DRD1", "hypothalamus") + theme(axis.title.x = element_blank())
b <- plotcandidateremoval(pitvsd, "PRL", "pituitary")+ theme(axis.title.x = element_blank())
c <- plotcandidateremoval(gonvsd, "CRHBP", "gonads")
